### 2023-02-15 paper-thesis/chapters/12-editor-services.tex
Meeting results
Minor writing changes
Add error annotations
Add chapter transitions

\section{Setup required}
To set up editor services, we followed the steps at \todo{cite} \url{https://spoofax.dev/spoofax-pie/develop/guide/static-semantics/code-completion/}. To be precise, we followed the following steps:
\begin{enumerate}
	\item Add \verb|tego-runtime {}| and \verb|code-completion {}| to the \verb|spoofax.cfg| file, to enable code completion.
	\item Add the following rules to the \verb|main.str2| file, to pre-process and post-process the AST for code completion.
	\begin{lstlisting}
rules
	downgrade-placeholders = downgrade-placeholders-MyLang
	upgrade-placeholders   = upgrade-placeholders-MyLang
	is-inj                 = is-MyLang-inj-cons
	pp-partial             = pp-partial-MyLang-string
	pre-analyze            = explicate-injections-MyLang
	post-analyze           = implicate-injections-MyLang
	\end{lstlisting}
	\item For each rule define a predicate that accepts a placeholder where a syntactic sort is permitted. For our language, those are the following:
	\begin{lstlisting}
expectBetaEq((_, Expr-Plhdr()), _).
expectBetaEq(_, (_, Expr-Plhdr())).
betaReduceHead((_, Expr-Plhdr())) = _.
betaReduce((_, Expr-Plhdr())) = _.
typeOfExpr(_, Expr-Plhdr()) = _.
programOk(Start-Plhdr()).
	\end{lstlisting}

@@

>>>>

\section{Setup required}
To set up editor services, we followed the steps in the Spoofax documentation \cite{semcomplet}. To be precise, we followed the following steps:
\begin{enumerate}
	\item Add \verb|tego-runtime {}| and \verb|code-completion {}| to the \verb|spoofax.cfg| file, to enable code completion.
	\item Add the following rules to the \verb|main.str2| file, to pre-process and post-process the AST for code completion.
	\begin{lstlisting}
rules
	downgrade-placeholders = downgrade-placeholders-MyLang
	upgrade-placeholders   = upgrade-placeholders-MyLang
	is-inj                 = is-MyLang-inj-cons
	pp-partial             = pp-partial-MyLang-string
	pre-analyze            = explicate-injections-MyLang
	post-analyze           = implicate-injections-MyLang
	\end{lstlisting}
	\item For each rule define a predicate that accepts a placeholder where a syntactic sort is permitted. For our language, those are the following:
	\begin{lstlisting}
expectBetaEq((_, Expr-Plhdr()), _).
expectBetaEq(_, (_, Expr-Plhdr())).
betaReduceHead((_, Expr-Plhdr())) = _.
betaReduce((_, Expr-Plhdr())) = _.
typeOfExpr(_, Expr-Plhdr()) = _.
programOk(Start-Plhdr()).
	\end{lstlisting}

@@

We have now discussed semantic code completion. Next, we will compare the implementation with one in Haskell (chapter~\ref{ch:comp-haskell}) and LambdaPi (chapter~\ref{ch:comp-lambdapi}).

### 2023-02-15 paper-thesis/chapters/22-adding-datatypes.tex
Work on CoC chapter
Fix chapter references
Add chapter transitions

Another useful feature is support for inductive data types. An example of a simple inductive datatype is the \verb|Nat| type:
\begin{lstlisting}
data Nat : -> Type where
    zero : Nat,
    suc : Nat -> Nat;
\end{lstlisting}
We will design data types to have the same features as in Agda:
\begin{enumerate}
\item Data types may be recursive, that is the data type may take itself as one of the constructor arguments. This can be seen in the definition of \verb|Nat| above.
\item Parameters, which are datatypes that are polymorphic, and are required to be the same for all constructors, such as
\begin{lstlisting}
data Maybe (T : Type) : -> Type where
    None : Maybe T,
    Some : T -> Maybe T;
\end{lstlisting}
\item Indices, which are datatypes that are polymorphic, that may vary from constructor to constructor
\begin{lstlisting}
data Eq : (e1: Bool) (e2 : Bool) -> Type where
    refl : e : Bool -> Eq e e;
\end{lstlisting}
\item We can also combine parameters and indices, for example to create the generic \verb|Eq| type. Note that parameters and indices may also depend on earlier parameters and indices.
\begin{lstlisting}
data Eq (T : Type) : (e1 : T) (e2 : T) -> Type where
    refl : e : T -> Eq e e;
\end{lstlisting}
\item Data types must be strictly positive. This ensures that we cannot make non-terminating programs\footnote{After we add support for universes in chapter \ref{ch:universes}}. For example, the following data type is forbidden, because it refers to itself in a negative position. We will explain exactly how this check works in section \ref{ch:datatypes:positivity}.
\begin{lstlisting}
data Bad : -> Type where
	bad : (Bad -> Bool) -> Bad;
\end{lstlisting}
\end{enumerate}

@@

\section{Type-checking data type declarations}
\label{ch:datatypes:typechecking}

@@

\section{Type-checking eliminators}

@@

>>>>

Another useful feature is support for inductive data types. An example of a simple inductive datatype is the \verb|Nat| type:
\begin{lstlisting}
data Nat : -> Type where
    zero : Nat,
    suc : Nat -> Nat;
\end{lstlisting}
We will design data types to have the same features as in Agda:
\begin{enumerate}
\item Data types may be recursive, that is the data type may take itself as one of the constructor arguments. This can be seen in the definition of \verb|Nat| above.
\item Parameters, which are datatypes that are polymorphic, and are required to be the same for all constructors, such as
\begin{lstlisting}
data Maybe (T : Type) : -> Type where
    None : Maybe T,
    Some : T -> Maybe T;
\end{lstlisting}
\item Indices, which are datatypes that are polymorphic, that may vary from constructor to constructor
\begin{lstlisting}
data Eq : (e1: Bool) (e2 : Bool) -> Type where
    refl : e : Bool -> Eq e e;
\end{lstlisting}
\item We can also combine parameters and indices, for example to create the generic \verb|Eq| type. Note that parameters and indices may also depend on earlier parameters and indices.
\begin{lstlisting}
data Eq (T : Type) : (e1 : T) (e2 : T) -> Type where
    refl : e : T -> Eq e e;
\end{lstlisting}
\item Data types must be strictly positive. This ensures that we cannot make non-terminating programs\footnote{After we add support for universes in chapter~\ref{ch:universes}}. For example, the following data type is forbidden, because it refers to itself in a negative position. We will explain exactly how this check works in section \ref{ch:datatypes:positivity}.
\begin{lstlisting}
data Bad : -> Type where
	bad : (Bad -> Bool) -> Bad;
\end{lstlisting}
\end{enumerate}

@@

\section{Type-checking Data Type Declarations}
\label{ch:datatypes:typechecking}

@@

\section{Type-checking Eliminators}

@@

We have now shown how to add inductive datatypes to the implementation. Next, we will Universes (chapter~\ref{ch:universes}) to the implementation.

### 2023-02-15 paper-thesis/chapters/23-adding-universes.tex
Add chapter transitions

>>>>

We have now finished the implementation of the language. Next, we will discuss how well semantic code completion works with this implementation in chapter~\ref{chap:editor-services}. Then we will compare the implementation with one in Haskell (chapter~\ref{ch:comp-haskell}) and LambdaPi (chapter~\ref{ch:comp-lambdapi}).
